Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-13 doc/report/msc-report.tex

\chapter{Analysis}

@@

\section{Task}

Mention data sets here.

\subsection{Import}

\subsection{Store}

\subsection{Export}

>>>>

\chapter{System Analysis}

\chapterquote{Methods and means cannot be separated from the ultimate aim.}{Emma
Goldman}

\section*{Introduction}

This chapter will discuss and analyze the system that is to be built. First the
function wanted from the system is presented; while discussing the input and
output characteristics, and possible challenges. Secondly, a rough sketch and
discussion of the system's different parts will be presented. Thirdly, the
requirements of the system will be featured.

Throughout this chapter distinction is made between the developer: who develops
the system, and the user: who uses the product which the developer
creates. Although, in reality, they might be the same person.

\section{System function}

Last.fm has a central statistics and metrics tool built on a relational
database. It is possible to import analyzed data to this tool; the data is
presented to the user as graphs. Drawbacks of the system include problems with
new data source additions: they can sometime not be loaded because they are to
big. If a user wants to perform a custom query, manual intervention by the
developer is necessary. There is no way for the system to handle custom queries.

The data source is typically a log file; it is often a large file, up to
gigabytes in size, of structured data. The data is structured the same way on
each line, this makes writing a program for analyzing the entire data set as
simple as analyzing a single line.

Reporting the analyzed is the data is the final step. This is what the user is
interested in: Viewing the analyzed data. The tool at Last.fm presents the user
with a dashboard that has a selection of graphs on primary metrics; it also has
a list of all available graphs. All the graphs are pre-aggregated, no support
exist for custom queries. Enabling the user to query the aggregated data---for
instance viewing the most popular web browser in a country---would open
possibilities to explore the data. During exploration interesting details could
possibly be discovered; details which may be excluded from the pre-rendered
aggregates.

Since the data sources at Last.fm are time-stamped logs, the constraint to only
handle, time-series was brought upon the system. This decision was reached
mainly to keep the implementation simple.

\subsection*{System parts}

The following three components form the outline of the system: importing,
storing, and exporting data.

At Last.fm there are several different facilities which produce logs, among
them: web servers, scrobbling\footnote{Last.fm users can submit what music they
listen to, this is called scrobbling.}, radio etc. This data reside on a
commodity hardware cluster used for parallel computation. The system needs to be
able to easily import data from this source.

% storage.
Because the data sources are typically large, the storage need to be efficient.
Storing the analyzed data can be done in different ways, for instance storing
the raw data or just aggregates. Storing the raw data is very expensive, since
it would basically copy the data to the data store. Because of this the decision
was made to only store aggregates.

There are several different ways of modelling multidimensional data.

TODO: How to discuss modelling of data in a general sense, without mentioning
MapReduce and BigTable?

% export.
Due to the several different data sources, there exist several different types
of applications. Therefor the system does not implement any user interface, this
is left for the developer to implement.

\section{System requirements}

The four key requirements of the system are: ease of use, speed, scalability,
and being able to fit in the existing software and hardware eco-system.

% ease of use.
The system should have an easy facility for importing and analyzing new data
sources, while still being fast and efficient. A simple script---or tweak of an
already existing data import script---should be enough for importing and
analyzing a new data source.

% speed.
Time is of essence when exporting data to the user. The user should be able to
explore the data at mouse-click rate, not having to wait for computation. This
is usually solved by pre-rendering the data which can be viewed or queried.

Importing data requires computing power, which is a limited resource. This calls
for the need to make the import phase efficient. Using a desktop workstation to
process several gigabytes of data is simply impossible; it might require more
time than the actual worth of data it is importing---for instance taking more
than one day to import a days worth of data. There are a number of ways to solve
this, among them: more powerful machines and parallelizing.

Acquiring more powerful machines quickly becomes expensive. Even if more
powerful machines are acquired they might not have enough power for processing
several gigabytes of data: the data to be processed might not fit in the working
memory.

Another technique is parallelizing, using a cluster of commodity hardware
machines. Using commodity hardware for computation significantly reduces the
cost, since commodity hardware is cheap; thusly preferred in most situations.

% scaling.
In order for a computational setup, cluster or otherwise, to be efficient it
needs to scale well, this means adding more hardware should add as close as
possible to the hardware's potential addition to the computational power.

% fit in existing eco system.
The system needs to co-exist and integrate seamlessly with the already existing
hardware and software eco-system at Last.fm. Thus, building on top of the
already existing software a natural choice.

\section*{Summary}

This chapter has analyzed and discussed the function and requirements of the
system to be built.

The system function part discussed the system's goals and parts. The goals are
to have a system which enables the developer to easily import and analyze new
data sources and to be fast enough to serve the analyzed data at mouse-click
rate to the user. The system's three core parts---importing, storing, and
extracting data---have also been discussed.

Furthermore the requirements on the system were discussed; the four key
requirements are: ease of use, speed, scaling, and fitting into the existing
hardware and software eco-system.

@@
